### 2023-01-01 documentation/paper.tex
FEATURE: Add section for approach and experiments

% ---------------------------------------------------------------------------
% Author guideline and sample document for EG publication using LaTeX2e input
% D.Fellner, v1.13, Nov 13, 2007

\documentclass{egpubl}

% --- for  Annual CONFERENCE
\VisualComputing % VisualComputing (added by U. Schwanecke, 2021)
%\ThreeDAnimation % 3DAnimation (added by U. Schwanecke, 2020)
%\Seminar % Specialist Seminar (added by U. Schwanecke, 2019)
%\ConferenceSubmission % uncomment for Conference submission
%\ConferencePaper      % uncomment for (final) Conference Paper
%\STAR                 % uncomment for STAR contribution
% \Tutorial             % uncomment for Tutorial contribution
% \ShortPresentation    % uncomment for (final) Short Conference Presentation
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to Computer Graphics Forum, special issue
% \SpecialIssuePaper         % uncomment for final version of Journal Paper, special issue
%
% --- for  EG Workshop Proceedings
% \WsSubmission    % uncomment for submission to EG Workshop
% \WsPaper         % uncomment for final version of EG Workshop contribution
%
%\electronicVersion % can be used both for the printed and electronic version

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------

\PrintedOrElectronic

\usepackage[pdftex]{graphicx} 
\usepackage{t1enc, dfadobe}
%\usepackage[T1]{fontenc}
\usepackage{egweblnk}
\usepackage{cite}

% For backwards compatibility to old LaTeX type font selection.
% Uncomment if your document adheres to LaTeX2e recommendations.
% \let\rm=\rmfamily    \let\sf=\sffamily    \let\tt=\ttfamily
% \let\it=\itshape     \let\sl=\slshape     \let\sc=\scshape
% \let\bf=\bfseries

% end of prologue

\usepackage{blindtext}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{dblfloatfix}
\usepackage{adjustbox}
\usepackage{mathtools}
\usepackage{todonotes}

% ---------------------------------------------------------------------
\captionsetup{labelfont=bf,textfont=it}
\title[Burst Images]%
      {Burst Images}

\author[Sascha Scheid]
    {\parbox{\textwidth}
        {\centering 
			Sascha Scheid
        }
        \\
    {\parbox{\textwidth}
        {\centering RheinMain University of Applied Sciences, Wiesbaden, Germany\\
       }
    }
}
% ------------------------------------------------------------------------

% if the Editors-in-Chief have given you the data, you may uncomment
% the following five lines and insert it here
%
% \volume{36}   % the volume in which the issue will be published;
% \issue{1}     % the issue number of the publication
% \pStartPage{1}      % set starting page

%-------------------------------------------------------------------------
\begin{document}

%-------------------------------------------------------------------------
% uncomment for using teaser (comment if you don't want a teaser)
\teaser{
    \hspace{\fill}
    \includegraphics[width=0.9\linewidth]{images/ModelbasedTrackingExample.png}
    \hspace{\fill}
    \centering
    \caption{Best teaser image you can provide \ldots}
    \label{fig:teaser}
}

%-------------------------------------------------------------------------
\maketitle

%-------------------------------------------------------------------------
\begin{abstract}
Some informative abstract \ldots
\blindtext
\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}

Burst photography is a technique in which a series of photographs is taken quickly in succession.
This can be useful in a variety of situations, such as capturing action or movement, 
or to create a sense of motion. In digital photography, burst images are stored as a 
sequence of image files, typically in a format such as JPEG or RAW.

High Dynamic Range (HDR) is a technique used to improve the dynamic range of an image, 
which is the range of luminance or brightness levels that can be captured in a photograph. 
The dynamic range of a scene can often be greater than what a camera is able to capture 
in a single image, resulting in lost detail in the highlights or shadows. 
HDR techniques can be used to extend the range of luminance in an image, 
resulting in more detail and a more realistic representation of the scene.

One way that burst photography and HDR techniques can be used together
is to capture a series of photographs with different exposures in a burst, and then combine 
the exposures into a single HDR image using software. This can be particularly useful in 
low-light situations, where the camera may struggle to capture a wide range of luminance 
levels in a single exposure. By capturing multiple exposures and combining them into an 
HDR image, it is possible to extend the dynamic range and capture more detail in both the 
highlights and shadows.

%-------------------------------------------------------------------------
\section{Related Work}
\label{sec:related_work}
Maybe a separate section about related work \ldots 
\blindtext

%-------------------------------------------------------------------------
\section{Approach}
\label{sec:approach}

The classic manual approach from digital photography is to take three images of the same scene: 
One overexposed, one underexposed, and one normally exposed. These are superimposed in 
post-processing so that the bright areas of the resulting image are not too bright and the 
dark areas are not too dark. This results in an image that not only contains many bright areas, 
but also many dark areas at the same time.

This classical approach will be contrasted with the method described in \ref{sec:related_work}. 
For this purpose, image bursts of the same scene are processed once according to the 
classical method and once according to the method described in \ref{sec:related_work}. To get a good 
comparison, images of stationary and moving objects are taken, both from a tripod and 
handheld.

%-------------------------------------------------------------------------
\section{Experiments}
\label{sec:Experiments}
\cite{Brooks2016git} has implemented the image processing pipeline described in \ref*{sec:related_work}. To do this, 
he uses Halide, a programming language embedded in C++ that allows efficient image processing 
pipelines to be implemented with little effort. Halide itself requires LLVM to compile. 
Unfortunately, the project has not been actively maintained for quite some time and 
since new versions of Halide and LLVM have been released in the meantime, some of which were 
not backwards compatible, the project could not be compiled out of the box. After many hours of 
trial and error, the project was able to compile with version 10 of both Halide and LLVM. 
To get a reproducible executable version of \cite{Brooks2016git}, a Docker image was created that 
contains the application and runs it on container startup.

%-------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusion}
Some famous last words \ldots
\blindtext

%-------------------------------------------------------------------------
\section{Acknowledgements}
\label{sec:acknowledgements}
Maybe you like to acknowledge \ldots

%-------------------------------------------------------------------------
%\bibliographystyle{eg-alpha}
\bibliographystyle{eg-alpha-doi}
\bibliography{paper-bib}

\end{document}